---
title: "Create Variables: 21-40"
author: "Joy"
date: "5/23/2021"
output: html_document
---
```{r package}
library(dplyr)
library(ggplot2)
library(corrplot)
library(fastDummies)
library(stats)
```

```{r data}
data <- read.csv('train.csv',header = TRUE)
dat <- cbind(data[,21:40], data[,81])
names(dat)[21] <- 'SalePrice'
str(dat)
```
```{r Correlation}
temp_df <- cor(na.omit(dat[, sapply(dat, is.numeric)]))
temp_df[abs(temp_df)<0.7] <- 0
corrplot(temp_df,type = "upper",
         number.cex = 0.5,method = 'square',tl.cex=0.6)
```

# 1 YearRemodAdd
YearRemodAdd : Remodel date
```{r YearRemodAdd}
str(dat[,1])
plot(table(dat[,1]))
# transform the data into the age of the remodeled
remodel_age <- 2021 - dat[,1]
f1 <- as.data.frame(remodel_age) # f stands for feature
colnames(f1) <- 'YearRemodAdd'
head(f1)
```

# 2 RoofStyle
RoofStyle: Type of roof  
       Flat	    Flat  
       Gable	  Gable  
       Gambrel	Gabrel (Barn)  
       Hip	    Hip  
       Mansard	Mansard  
       Shed	    Shed  
```{r RoofStyle 1/2}
sum(is.na(dat[,2])) # no NAs
table(dat[,2]) # check distribution
# make into dummy vairables
temp_roof <- dummy_cols(dat, select_columns = "RoofStyle")[21:27]
cor(temp_roof)[1,]
corrplot(cor(temp_roof),type = "upper",
         number.cex = 0.5,method = 'square',tl.cex=0.6)
```
Since `Flat`, `Mansard`, `Shed`and `Gambrel` only account for less than 50 instances, we will combine them into `Others`.  
```{r combine RoofStyle 2/2}
# combine
f2 <- ifelse(dat[,2] %in% c('Flat', 'Mansard', 'Shed','Gambrel'),
                            'Others', dat[,2])
table(f2)
f2 <- as.data.frame(as.factor(f2))
names(f2) <- "RoofStyle"
head(f2)
```

# 3 RoofMatl
**exclude this variable**
RoofMatl: Roof material
       ClyTile	Clay or Tile
       CompShg	Standard (Composite) Shingle
       Membran	Membrane
       Metal	Metal
       Roll	Roll
       Tar&Grv	Gravel & Tar
       WdShake	Wood Shakes
       WdShngl	Wood Shingles
```{r RoofMatl}
sum(is.na(dat[,3])) # no NAs
table(dat[,3])
```
# 4 Exterior1st
Exterior1st: Exterior covering on house
       AsbShng	Asbestos Shingles
       AsphShn	Asphalt Shingles
       BrkComm	Brick Common
       BrkFace	Brick Face
       CBlock	Cinder Block
       CemntBd	Cement Board
       HdBoard	Hard Board
       ImStucc	Imitation Stucco
       MetalSd	Metal Siding
       Other	Other
       Plywood	Plywood
       PreCast	PreCast	
       Stone	Stone
       Stucco	Stucco
       VinylSd	Vinyl Siding
       Wd Sdng	Wood Siding
       WdShing	Wood Shingles
```{r Exterior1st}
sum(is.na(dat[,4])) # no NAs
table(dat[,4])
cor(dummy_cols(dat, select_columns = "Exterior1st")[21:36])[1,]

# anything less than 61, will be combined to "others"

f4 <- ifelse(dat[,4] %in% c('AsbShng', 'AsphShn', 'BrkComm','BrkFace',
                                 'CBlock','ImStucc', 'Stone', 'Stucco', 'WdShing'),
                  'Others', dat[,4])
table(f4)

f4 <- as.data.frame(as.factor(f4))
names(f4) <- "Exterior1st"
head(f4)
```

# 5 Exterior2nd
**exclude this variable**
Exterior2nd: Exterior covering on house (if more than one material)
```{r Exterior2nd}
sum(is.na(dat[,5])) # no NAs
table(dat[,5])
cor(dummy_cols(dat, select_columns = "Exterior2nd")[21:37])[1,]
```

# 6 MasVnrType
MasVnrType: Masonry veneer type
       BrkCmn	Brick Common
       BrkFace	Brick Face
       CBlock	Cinder Block
       None	None
       Stone	Stone
```{r MasVnrType 1/2}
sum(is.na(dat[,6])) # 8 NAs
# categorizing those as none
dat[is.na(dat[,6])==1,]$MasVnrType <- 'None'
table(dat[,6])
cor(dummy_cols(dat, select_columns = "MasVnrType")[21:25])[1,]

f6 <- as.data.frame(as.factor(dat[,6]))
names(f6) <- "MasVnrType"
head(f6)
```
# 7 MasVnrArea
MasVnrArea: Masonry veneer area in square feet
```{r MasVnrArea}
# histogram
ggplot(dat, aes(x = MasVnrArea)) + geom_histogram(bins = 100) +
  xlab('Masonry veneer area in square feet')

# histogram without 0 sqft
ggplot(dat[dat$MasVnrArea > 1,], aes(x = MasVnrArea)) + 
  geom_histogram(bins = 100) +
  xlab('Masonry veneer area in square feet')

# scatter plot with sales price
ggplot(dat, aes(x = MasVnrArea, y = SalePrice)) + geom_point(alpha = .3) +
  xlab('Masonry veneer area in square feet') +
  ylab('Sales Price')
```
```{r MasVnrArea fill in NAs}
sum(is.na(dat[,7])) # 8 NAs
# fill in 0 for NAs
dat[is.na(dat[,7])==1,]$MasVnrArea <- 0
f7 <- as.data.frame(dat[,7])
colnames(f7) <- 'MasVnrArea'
```

# 8 ExterQual
ExterQual: Evaluates the quality of the material on the exterior  
       Ex	Excellent  
       Gd	Good  
       TA	Average/Typical  
       Fa	Fair  
       Po	Poor  
```{r ExterQual}
sum(is.na(dat[,8])) # no NAs
table(dat[,8])
cor(dummy_cols(dat, select_columns = "ExterQual")[21:25])[1,]

f8 <- as.data.frame(as.factor(dat[,8]))
names(f8) <- "ExterQual"
head(f8)
```

# 9 ExterCond
ExterCond: Evaluates the present condition of the material on the exterior  
       Ex	Excellent
       Gd	Good
       TA	Average/Typical
       Fa	Fair
       Po	Poor
```{r ExterCond}
sum(is.na(dat[,9])) # no NAs
table(dat[,9])
cor(dummy_cols(dat, select_columns = "ExterCond")[21:26])[1,]

# combining fair and poor
f9 <- ifelse(dat[,9]=='Po', 'Fa', dat[,9])
table(f9)

f9 <- as.data.frame(as.factor(dat[,9]))
names(f9) <- "ExterCond"
head(f9)
```

# 10 Foundation
Foundation: Type of foundation
       BrkTil	Brick & Tile
       CBlock	Cinder Block
       PConc	Poured Concrete	
       Slab	Slab
       Stone	Stone
       Wood	Wood
```{r Foundation}
sum(is.na(dat[,10])) # no NAs
table(dat[,10])
cor(dummy_cols(dat, select_columns = "Foundation")[21:27])[1,]

# combine vairables with instances less than 50 as "Others"
f10 <- ifelse(dat[,10] %in% c('Slab', 'Stone', 'Wood'), "Others",dat[,10])
table(f10)

f10 <- as.data.frame(as.factor(f10))
names(f10) <- "Foundation"
head(f10)
```

# 11 BsmtQual
# 12 BsmtCond
BsmtQual: Evaluates the height of the basement
       Ex	Excellent (100+ inches)  	   5
       Gd	Good (90-99 inches)          4
       TA	Typical (80-89 inches)       3
       Fa	Fair (70-79 inches)          2
       Po	Poor (<70 inches             1
       NA	No Basement                  0
# 13 BsmtExposure
       Gd	Good Exposure                5
       Av	Average Exposure 	           3
       Mn	Mimimum Exposure             1
       No	No Exposure                  0
       NA	No                           0
# 14 BsmtFinType1
BsmtFinType1: Rating of basement finished area  
       GLQ	Good Living Quarters                5
       ALQ	Average Living Quarters             4
       BLQ	Below Average Living Quarters	      3
       Rec	Average Rec Room                    2
       LwQ	Low Quality                         1
       Unf	Unfinshed                           0
       NA	No Basement                           0
# 16 BsmtFinType2
```{r 11 BsmtQual}
sum(is.na(dat[,11])) # 37 no Basements named "NA"
# change NA to NB meaning no basements
dat[is.na(dat[,11]) ==1,]$BsmtQual <- 'NB'
table(dat[,11])
f11 <- as.data.frame(ifelse(dat$BsmtQual == 'NB', 0 ,1))
names(f11) <- 'HasBsmt'

# assign score to the bsmt quality
bsmtscore11 <- as.data.frame(ifelse(dat$BsmtQual == 'Ex', 5,
                                    ifelse(dat$BsmtQual == 'Gd', 4,
                                           ifelse(dat$BsmtQual == 'TA', 3,
                                                  ifelse(dat$BsmtQual == 'Fa', 2,
                                                         ifelse(dat$BsmtQual =='Po',
                                                                1,0))))))

```

```{r 12 BsmtCond}
sum(is.na(dat[,12])) # 37 no Basements named "NA"
# change NA to NB meaning no basements
dat[is.na(dat[,12]) ==1,]$BsmtCond <- 'NB'
table(dat[,12])
cor(dummy_cols(dat, select_columns = "BsmtCond")[21:26])[1,]

bsmtscore12 <- as.data.frame(ifelse(dat$BsmtCond == 'Ex', 5,
                                    ifelse(dat$BsmtCond == 'Gd', 4,
                                           ifelse(dat$BsmtCond == 'TA', 3,
                                                  ifelse(dat$BsmtCond == 'Fa', 2,
                                                         ifelse(dat$BsmtCond =='Po',
                                                                1,0))))))
```

```{r 13 BsmtExposure}
sum(is.na(dat[,13])) # 38 "NA" but only 37 no basement
# change NA to NB, including the one without data
dat[is.na(dat[,13]) ==1,]$BsmtExposure <- 'NB'
table(dat[,13])
cor(dummy_cols(dat, select_columns = "BsmtExposure")[21:26])[1,]

bsmtscore13 <- as.data.frame(
  ifelse(dat$BsmtExposure == 'Gd', 5,
         ifelse(dat$BsmtExposure == 'Av', 3,
                ifelse(dat$BsmtExposure == 'Mn', 1,0))))
```
  
```{r 14 BsmtFinType1}
sum(is.na(dat[,14])) # 37 no basement
# change NA to NB
dat[is.na(dat[,14]) ==1,]$BsmtFinType1 <- 'NB'
table(dat[,14])
cor(dummy_cols(dat, select_columns = "BsmtFinType1")[21:28])[1,]

bsmtscore14 <- as.data.frame(
  ifelse(dat$BsmtCond == 'GLQ', 5,
         ifelse(dat$BsmtFinType1 == 'ALQ', 4,
                ifelse(dat$BsmtFinType1 == 'BLQ', 3,
                       ifelse(dat$BsmtFinType1 == 'Rec', 2,
                              ifelse(dat$BsmtFinType1 =='LwQ',1,0))))))
```

```{r 16 BsmtFinType2}
sum(is.na(dat[,16])) # 38 no basement
# change NA to NB
dat[is.na(dat[,16]) ==1,]$BsmtFinType2 <- 'NB'
table(dat[,16])
cor(dummy_cols(dat, select_columns = "BsmtFinType2")[21:28])[1,]

bsmtscore16 <- as.data.frame(
  ifelse(dat$BsmtCond == 'GLQ', 5,
         ifelse(dat$BsmtFinType2 == 'ALQ', 4,
                ifelse(dat$BsmtFinType2 == 'BLQ', 3,
                       ifelse(dat$BsmtFinType2 == 'Rec', 2,
                              ifelse(dat$BsmtFinType2 =='LwQ',1,0))))))
```
# Creating Average Basement Score
```{r Bsmt Score}
f12 <- (bsmtscore11 + bsmtscore12 + bsmtscore13 + bsmtscore14 +
                   bsmtscore16)/5
names(f12) <- "AvgBsmtScore"
head(f12)
```

# 15 BsmtFinSF1
BsmtFinSF1: Type 1 finished square feet
``` {r BsmtFinSF1}
# scatter plot with sales price
ggplot(dat, aes(x = BsmtFinSF1, y = SalePrice)) + geom_point(alpha = .3) +
  xlab('Basement Type 1 finished square feet') +
  ylab('Sales Price')
  
# histogram
ggplot(dat, aes(x = BsmtFinSF1)) + geom_histogram(bins = 100) +
    xlab('Basement Type 1 finished square feet')

# histogram without 0 sqft
ggplot(dat[dat$BsmtFinSF1 > 1,], aes(x = BsmtFinSF1)) + 
  geom_histogram(bins = 100) +
  xlab('Basement Type 1 finished square feet')
```
# 17 BsmtFinSF2
BsmtFinSF2: Type 2 finished square feet
```{r BsmtFinSF2}
# scatter plot with sales price
ggplot(dat, aes(x = BsmtFinSF2, y = SalePrice)) + geom_point(alpha = .3) +
  xlab('Basement Type 2 finished square feet') +
  ylab('Sales Price')
  
# histogram
ggplot(dat, aes(x = BsmtFinSF2)) + geom_histogram(bins = 100) +
    xlab('Basement Type 2 finished square feet')

# histogram without 0 sqft
ggplot(dat[dat$BsmtFinSF2 > 1,], aes(x = BsmtFinSF2)) + 
  geom_histogram(bins = 100) +
  xlab('Basement Type 2 finished square feet')
```

# 18 BsmtUnfSF
BsmtUnfSF: Unfinished square feet of basement area
```{r BsmtUnfSF}
# scatter plot with sales price
ggplot(dat, aes(x = BsmtUnfSF, y = SalePrice)) + geom_point(alpha = .3) +
  xlab('Unfinished square feet of basement area') +
  ylab('Sales Price')
  
# histogram
ggplot(dat, aes(x = BsmtUnfSF)) + geom_histogram(bins = 100) +
  xlab('Unfinished square feet of basement area')

# histogram without 0 sqft
ggplot(dat[dat$BsmtUnfSF > 1,], aes(x = BsmtUnfSF)) + 
  geom_histogram(bins = 100) +
  xlab('Unfinished square feet of basement area')
```
# 19 TotalBsmtSF
TotalBsmtSF: Total square feet of basement area
```{r TotalBsmtSF}
options(scipen=999)
# scatter plot with sales price
ggplot(dat, aes(x = TotalBsmtSF, y = SalePrice)) + geom_point(alpha = .3) +
  xlab('Total square feet of basement area') +
  ylab('Sales Price')
  
# histogram
ggplot(dat, aes(x = TotalBsmtSF)) + geom_histogram(bins = 100) +
  xlab('Total square feet of basement area')

f19 <- as.data.frame(dat[,19])
colnames(f19) <- 'TotalBsmtSF'
```

# 20 Heating
Heating: Type of heating
       Floor	Floor Furnace
       GasA	Gas forced warm air furnace
       GasW	Gas hot water or steam heat
       Grav	Gravity furnace	
       OthW	Hot water or steam heat other than gas
       Wall	Wall furnace
```{r Heating}
sum(is.na(dat[,20])) # no NAs
table(dat[,20])
cor(dummy_cols(dat, select_columns = "Heating")[21:27])[1,]
# combining everything other than GasA to "Others"
f20 <- ifelse(dat[,20]!= 'GasA', 'OthrHeating', 'GasA')
table(f20)

f20 <- as.data.frame(as.factor(f20))
names(f20) <- "Heating"
head(f20)
```

# Creating Extra Variables
## 21 PctFinBsmt
Percentage of Finished Basement
```{r PctFinBsmt}
PctFinBsmt <- (dat[,15] + dat[,17])/dat[,19]
f21 <- as.data.frame(PctFinBsmt)
colnames(f21) <- 'PctFinBsmt'
f21[is.na(f21),] <- '0' # changing NAs to 0 for no basements
```

# Combine all dataframes
```{r cbind}
final <- cbind(dat$SalePrice, f1, f2, f4, f6, f7, f8, 
               f9, f10, f11, f12, f19, f20, f21)
names(final)[1] <- "SalePrice"
```
## Exporting Final File
```{r Export}
write.csv(final, '/Users/joychen/OneDrive - UC Irvine/288Predictive/course_project/house-prices-data/final_train.csv')
```
